---
title: "Geocoded aid and NGO crackdown"
format: 
  gfm: 
    wrap: preserve
---

<!-- README.md is generated from README.qmd. Please edit that file -->

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-",
  fig.retina = 3
)

library(emoji)

# manuscript_target <- targets::tar_read(manuscript)

# yaml_metadata <- rmarkdown::yaml_front_matter(here::here("manuscript", "manuscript.qmd"))
# title <- yaml_metadata$title
# abstract <- yaml_metadata$abstract

emoji_out <- emoji_glue(":lock::turtle:")
```


## `r emoji_out`: Note on "tight tortoise" project name

Because project titles change all the time with revisions, rewriting, and peer review, we used [{codename}](http://svmiller.com/codename/) to generate an [Ubuntu-style](https://wiki.ubuntu.com/DevelopmentCodeNames) internal-to-us project name that won't change.

```{r}
#| label: generate-code-name
#| eval: false

library(codename)
codename_message()
#> code name generated by {codename} v.0.5.0. R version 4.5.0 (2025-04-11).

codename(type = "ubuntu", seed = "geocoded aid ngo laws")
#> [1] "tight tortoise"
```


## Licenses

**Text and figures:** All prose and images are licensed under Creative Commons ([CC-BY-4.0](http://creativecommons.org/licenses/by/4.0/)).

**Code:** All code is licensed under the [MIT License](LICENSE.md).


## Contributions and Code of Conduct

We welcome contributions from everyone. Before you get started, please see our [contributor guidelines](CONTRIBUTING.md). Please note that this project is released with a [Contributor Code of Conduct](https://contributor-covenant.org/version/2/0/CODE_OF_CONDUCT.html). By contributing to this project, you agree to abide by its terms.
