---
title: "Civil Society Crackdowns and the Spatial Redistribution of Foreign Aid"
format: 
  gfm: 
    wrap: preserve
    toc: true
---

<!-- README.md is generated from README.qmd. Please edit that file -->

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-",
  fig.retina = 3
)

library(emoji)

# manuscript_target <- targets::tar_read(manuscript)

# yaml_metadata <- rmarkdown::yaml_front_matter(here::here("manuscript", "manuscript.qmd"))
# title <- yaml_metadata$title
# abstract <- yaml_metadata$abstract

emoji_out <- emoji_glue(":lock::turtle:")
```

---

[Andrew Heiss](https://www.andrewheiss.com) • Georgia State University  
[Suparna Chaudhry](http://www.suparnachaudhry.com/) • Lewis & Clark College

---

## Abstract

Anti-NGO laws have restricted civil society around the world and shape how official aid agencies allocate foreign aid. While existing research shows these restrictions reduce bilateral aid flows, little is known about their geographic effects. Using geocoded aid data for hundreds of thousands of projects from 1973-2020, combined with original civil society restriction data across 148 countries, we explore whether civil society restrictions lead to aid spillovers to neighboring countries and increased geographic dispersion of projects within restricted countries. Our findings reveal underexplored spatial dynamics of aid allocation and contribute to literature on the consequences of civil society restrictions.


## `r emoji_out`: Note on "tight tortoise" project name

Because project titles change all the time with revisions, rewriting, and peer review, we used [{codename}](http://svmiller.com/codename/) to generate an [Ubuntu-style](https://wiki.ubuntu.com/DevelopmentCodeNames) internal-to-us project name that won't change.

```{r}
#| label: generate-code-name
#| eval: false

library(codename)
codename_message()
#> code name generated by {codename} v.0.5.0. R version 4.5.0 (2025-04-11).

codename(type = "ubuntu", seed = "geocoded aid ngo laws")
#> [1] "tight tortoise"
```


## Licenses

**Text and figures:** All prose and images are licensed under Creative Commons ([CC-BY-4.0](http://creativecommons.org/licenses/by/4.0/)).

**Code:** All code is licensed under the [MIT License](LICENSE.md).


## Contributions and Code of Conduct

We welcome contributions from everyone. Before you get started, please see our [contributor guidelines](CONTRIBUTING.md). Please note that this project is released with a [Contributor Code of Conduct](https://contributor-covenant.org/version/2/0/CODE_OF_CONDUCT.html). By contributing to this project, you agree to abide by its terms.
